Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy

### 2020-01-23 manuscript/manuscript.Rmd
Oops, forgot to include this one too

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that structural factors such as domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Following the 2011 Egyptian uprising, just as domestic non-governmental organizations (NGOs) began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in its crackdown on NGOs. India, Bolivia, Hungary, Russia, and others have imposed funding restrictions on NGOs [@Carothers:2015; @DPR:2015]. Consequently, official aid flows as well as grants from large foundations have considerably decreased to NGOs in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018].

In response to this dramatic shift in the NGO funding landscape, organizations have increasingly turned to *foreign* individual donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. This is because donations from *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in the Global South often cannot support local NGOs, may channel funds to groups working on non-contentious issues such as health and education [@DPR:2015; @RPC:2016; @Brechenmacher:2017], or may be deterred by dismal tax incentives, threats to take away business permits, or a general fear of retribution [@HudsonInstitute:2015]. Therefore, NGOs have increasingly sought out funds from *foreign* individual donors in countries with institutions more amenable to philanthropy. By 2016, the amount directed towards civil society groups working in international affairs increased to \$22.03 billion [@givingusa:2017]. As Daniel Runde pointed out in a statement before a U.S. Senate subcommittee on multilateral institutions, “U.S. philanthropy is significant [for global philanthropic efforts] but it is mainly Ms. Smith and Mr. Johnson, [n]ot Mr. and Mrs. Gates” [@Runde:2017].

Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note, we argue that domestic political environments of NGO host countries can also serve as one such structural heuristic that can influence individual donor decisions. In particular, legal crackdowns on NGOs serve as a heuristic to donors of organizational deservingness and when donors are provided salient information about crackdown on NGOs and its impact on NGO activities, it can influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organizational-level heuristics like NGO issue area and its main funding sources.

This research note makes two important contributions. While recent work has looked at how state repression of NGOs affects official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how repression affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about structural determinants that influence philanthropy, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

@@

Substantial research on charitable-giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, little research has focused on the impact of structural factors on the choice to donate. Moreover, research on individual donor behavior has overwhelming looked at giving to local, rather than international causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore private philanthropic giving because individuals rely on signals and heuristics that differ from the benchmarks that large donors (whether official or private) rely on.

The decision to donate to an NGO is rarely rational. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourage or discourage overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010; @SzperPrakash:2011]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGOs, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

In addition to the above factors that signal an NGO's trustworthiness and deservingness, we argue that structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use permissive regulations to benignly routinize the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors perceive legal crackdowns differently from organizational-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. On the other hand, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged NGOs, even if donors are not completely satisfied with some organizational-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated:

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor decisions (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs both when organizational issue areas (**I**) are in tension in with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics thus confound the causal effect of crackdowns on donor decisions and must be accounted for in the analysis—issue areas and funding sources simultaneously influence the likelihood of crackdown while also serving as organizational heuristics that influence donor perceptions of NGO deservingness.

@@

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017; @Swiney:2019]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. Thus, donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on political causes.

We hypothesize that donors will be more likely to donate to less contentious NGOs because these are seen as uncontroversial, apolitical, and directed towards the most deserving of need, and are more capable of producing short-term quantifiable results. We use humanitarian NGOs as a proxy for NGOs working on non-contentious issues and human rights NGOs as a proxy for contentious issues.

@@

The final heuristic that can influence individual donors that we examine is the *source of NGO funding*. Government grants to NGOs can often crowd out private funding for nonprofit organizations [@AndreoniPayne:2003; @Brooks:2000], though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the donor government even if it were in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help.

As funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015], organizational funding again confounds the effect of anti-NGO crackdowns on donor behavior (see @fig:causal-path). As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Individual donors may be less likely to donate to government-funded NGOs that face restrictions, since legal trouble with the organization's primary funder could fuel private donor mistrust (i.e. perhaps the NGO did something to deserve crackdown). Donors thus may be more likely to donate more to NGOs that rely on private funding.

@@

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the countries an NGO works in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?". For the sake of this experiment, this ambiguous definition is not an issue. The heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if they don't know the exact subfield of human rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic.

@@

To measure the effect of these signals on the propensity and amount of individual donations, we use an experiment to vary the different organizational and structural signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs optimize and tailor their fundraising strategies for private individuals.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donation, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

Without considering issue or funding, we find that respondents are not more likely to donate to legally besieged international NGOs (H~1a~). As seen in @fig:likely-diffs(A), those exposed to the crackdown condition tend to have a slightly higher probability of donating (`r percent_1(filter(likely, title == "Crackdown")$theta_2)`) than those in the control condition (`r percent_1(filter(likely, title == "Crackdown")$theta_1)`), but the difference is not significant (`r filter(likely, title == "Crackdown")$delta`; `r filter(likely, title == "Crackdown")$p_neq0`). We find similar results when looking at the isolated effects of issue area and funding. Both differences are positive, but the posterior probability of being larger than zero is low (`r filter(likely, title == "Issue")$delta`; `r filter(likely, title == "Issue")$p_neq0`; `r filter(likely, title == "Funding")$delta`; `r filter(likely, title == "Funding")$p_neq0`). We thus do not initially find evidence that crackdowns (or issue or funding) alone increase the likelihood of donating.

@@

@fig:likely-diffs(B–C) shows the difference in donation likelihood for both crackdown conditions across NGO issue areas (H~2a~) and funding sources (H~3a~). Crackdown has almost no effect on the likelihood of donating to human rights NGOs (`r filter(likely, title == "Human rights")$p_neq0`). We hypothesized that the positive crackdown effect would be stronger for humanitarian NGOs, and while crackdown does have a positive effect under the humanitarian frame, the probability this difference is greater than zero is below our 95% threshold (`r filter(likely, title == "Humanitarian assistance")$p_neq0`). We do find some evidence for H~3a~, however. Crackdown has little effect on the likelihood of donating to NGOs receiving government funding (`r filter(likely, title == "Government")$p_neq0`), but it does increase the likelihood of donating to privately funded NGOs by `r percent_0(filter(likely, title == "Private")$pct_change)` (`r filter(likely, title == "Private")$delta`; `r filter(likely, title == "Private")$p_neq0`).

While the crackdown frame has no effect on the probability of donating to human rights NGOs alone, conditioning this finding on the source of NGO funding reveals competing trends. @fig:likely-diffs(D) shows the difference in donation likelihood across all experimental conditions. When donors know that a human rights organization is primarily government funded, they are more likely to donate in the *absence* of crackdowns than when the NGO faces legal difficulties (`r filter(likely, title == "Human rights | Government")$delta`; `r filter(likely, title == "Human rights | Government")$p_neq0`). In contrast, when donors know that an organization is privately funded, they are more likely to donate when there *is* a legal crackdown than when the NGO faces no legal hurdles (`r filter(likely, title == "Human rights | Private")$delta`; `r filter(likely, title == "Human rights | Private")$p_neq0`). For humanitarian assistance NGOs, however, legal crackdowns do not have a significant effect on the likelihood of donation when nesting funding and issue frames.

These results give partial support to H~3a~. The source of funding has little influence on the propensity to donate to humanitarian NGOs, but on average, donors are substantially more likely to give to privately-funded human rights NGOs than government-funded organizations. However, the effect of crackdowns reverses across conditions, as donors appear to both punish government-funded human rights NGOs that face legal difficulties and rally behind privately-funded NGOs facing those same hurdles. Respondents described their reasoning for the former as "[i]f the non-profit works in countries where non-profit activities are heavily restricted there's a good amount of waste and shortfalls in their activities and ability to assist others."[^3154467] Respondents also explained their motivation for increased support to privately funded NGOs as "[t]hey [IRC] are doing good work in countries where it is tough for groups like them to operate and they need all the help they can get."[^3025921] Individual donors thus seem to be more willing to support besieged human rights organizations when they are unencumbered by government funds.

@@

Crackdowns do not substantially influence donors' *likelihood* to donate, but they do increase the *amount* of money that respondents are willing to contribute (see @fig:amount-diffs(A)). Informing participants that IRC faces legal hurdles abroad increased donations by `r md_dollar(filter(amount, title == "Crackdown")$mu_diff)` on average, a `r percent_0(filter(amount, title == "Crackdown")$pct_change)` increase from the no crackdown condition (`r filter(amount, title == "Crackdown")$p_neq0`). This provides good evidence for H~1b~—donors appear to give more money to besieged international NGOs. In contrast, neither issue area nor funding have an effect on the amount donated (`r filter(amount, title == "Issue")$delta`; `r filter(amount, title == "Issue")$p_neq0`; `r filter(amount, title == "Funding")$delta`; `r filter(amount, title == "Funding")$p_neq0`), indicating that crackdown is a more powerful heuristic for determining private donations.

@@

Combining issue areas and funding sources adds more texture to these results. We previously found that donors seem to (1) punish government-funded human rights NGOs and (2) reward privately-funded human rights NGOs when they face legal issues. This finding also applies to the amount donors give. Respondents donated `r md_dollar(filter(amount, title == "Human rights | Private")$mu_diff)` more to besieged privately-funded human rights NGOs compared to the no crackdown control condition (`r filter(amount, title == "Human rights | Private")$p_neq0`), increasing their donations by `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)`. Respondents explained that they chose to give additional money to privately-funded human rights NGOs precisely because "the country limits non-profits."[^13685] The punishment mechanism found previously does not translate to the amount donated—while there is a decrease of `r md_dollar(abs(filter(amount, title == "Human rights | Government")$mu_diff))` in average donations to government-funded human rights NGOs facing crackdown, the chance this difference is less than zero is lower than our threshold (`r filter(amount, title == "Human rights | Government")$p_neq0`).

>>>>

---
title: "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  email: suparna.chaudhry@cnu.edu
  url: http://www.suparnachaudhry.com/
  twitter: SuparnaChaudhry
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
  twitter: andrewheiss
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. With reduced official aid and foundation funding available for NGOs, and an absence of philanthropic culture in many countries across the Global South, various NGOs are seeking funds from individual private donors in Western countries. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? While existing research has shown that donors rely on various heuristics from social networks and organizational-level attributes, we argue that domestic political environments of NGO host countries can also serve as an important structural-level heuristic that can influence donor behavior. Using a survey experiment, we find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We discuss the implications of our results on the sustainability of organizations working in repressive countries.
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of Political Science, Christopher Newport University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of Political Science, Christopher Newport University, Newport News, VA 23606, 757-594-7821. Contact: suparna.chaudhry@cnu.edu
word-count: "4,081"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Following the 2011 Egyptian uprising, just as domestic non-governmental organizations (NGOs) began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in this endeavor-India, Bolivia, Hungary, Russia, and others have also imposed funding restrictions on NGOs [@Carothers:2015; @DPR:2015]. Consequently, official aid flows as well as grants from large foundations have considerably decreased to NGOs in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018].

In response to this shift in the NGO funding landscape, organizations have increasingly turned to *foreign* individual donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. This is because donations from *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in the Global South often cannot support local NGOs, may prefer channeling funds to groups working on innocuous issues such as health and education over more contentious issues [@DPR:2015; @RPC:2016; @Brechenmacher:2017], or may be deterred by dismal tax incentives or a general fear of retribution [@HudsonInstitute:2015]. Therefore, NGOs have increasingly sought out funds from *foreign* individual donors in countries with institutions more amenable to philanthropy. This amount is not insignificant-by 2016, the amount directed towards civil society groups working in international affairs had increased to \$22.03 billion [@givingusa:2017].

In the face of increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate? Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974]. While much of this research has focused on heuristics from social networks and organizational practices [@CrosonShang:2011; @Tremblay-BoirePrakash:2017], in this research note, we argue that domestic political environments of NGO host countries can serve as one such structural heuristic that can also influence individual donor decisions. When donors are provided salient information about crackdown on NGOs and its impact on their activities, it can serve as a heuristic of organizational deservingness, and consequently influence donor preferences and behavior. We use a survey experiment to measure the effect of crackdowns on stated donor behavior and find that while crackdowns do not directly influence the likelihood of donation on their own, respondents indicated willingness to donate substantially more to legally restricted NGOs. This effect persists when mediated by other organizational-level heuristics like NGO issue area and its main funding sources.

This research note makes two important contributions. While recent work has looked at how crackdowns on NGOs affect official aid [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how this affects private aid. Further, while there is a robust existing literature on the individual-level determinants of philanthropy, far less is known about its structural-level determinants, particularly towards international causes, and how these factors interact with organizational attributes to signal NGO deservingness.

@@

Substantial research on charitable-giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011], and more recently, the effect of social networks on philanthropy [@HerzogYang:2018]. However, research on individual donor behavior has overwhelming examined giving to local, rather than international causes. While most research on giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014], we explore private philanthropic giving because individuals rely on signals and heuristics that may differ from the benchmarks that large donors (whether official or private) rely on.

The decision to donate to an NGO is rarely rational. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. Similarly, the effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010; @SzperPrakash:2011]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009].

In addition to the above factors that signal an NGO's trustworthiness and deservingness, we argue that structural factors—including whether an organization works abroad and the domestic political environment of the group's host country—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between INGOs and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use laws to regulate the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors perceive legal crackdowns differently from organizational-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. On the other hand, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged NGOs, even if donors are not completely satisfied with some organizational-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated.

@@

@fig:causal-path represents this relationship more formally. Legal crackdowns (**C**) have a direct effect on donor decisions (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs both when organizational issue areas (**I**) are in tension with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics thus confound the causal effect of crackdowns on donor decisions and must be accounted for in the analysis—issue areas and funding sources simultaneously influence the likelihood of crackdown while also serving as organizational heuristics that influence donor perceptions of NGO deservingness.

@@

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. Thus, donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes. As such, we hypothesize that donors will be more likely to donate to less contentious NGOs because these are seen as uncontroversial, apolitical, directed towards the most deserving of need, and are more capable of producing short-term quantifiable results. We use humanitarian NGOs as a proxy for NGOs working on non-contentious issues and human rights NGOs as a proxy for contentious issues.

@@

The final heuristic that can influence individual donors that we examine is the *source of NGO funding*. Government grants to NGOs can often crowd out private funding for nonprofit organizations [@AndreoniPayne:2003; @Brooks:2000], though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help.

As funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015], organizational funding again confounds the effect of anti-NGO crackdowns on donor behavior (see @fig:causal-path). As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Individual donors may be less likely to donate to government-funded NGOs that face restrictions, since legal trouble with the organization's primary funder could fuel private donor mistrust (i.e. perhaps the NGO did something to deserve crackdown). Donors thus may be likely to donate more to NGOs that rely on private funding.

@@

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguous definition is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic.

@@

To measure the effect of structural and organizational-level heuristics on individual donor preferences, we use an experiment to vary the different signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs optimize and tailor their fundraising strategies for private individuals.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

Without considering issue or funding, we find that respondents are not more likely to donate to legally besieged INGOs (H~1a~). As seen in @fig:likely-diffs(A), those exposed to the crackdown condition tend to have a slightly higher probability of donating (`r percent_1(filter(likely, title == "Crackdown")$theta_2)`) than those in the control condition (`r percent_1(filter(likely, title == "Crackdown")$theta_1)`), but the difference is not significant (`r filter(likely, title == "Crackdown")$delta`; `r filter(likely, title == "Crackdown")$p_neq0`). We find similar results when looking at the isolated effects of issue area and funding. Both differences are positive, but the posterior probability of being larger than zero is low (`r filter(likely, title == "Issue")$delta`; `r filter(likely, title == "Issue")$p_neq0`; `r filter(likely, title == "Funding")$delta`; `r filter(likely, title == "Funding")$p_neq0`). We thus do not initially find evidence that crackdowns (or issue or funding) alone increase the likelihood of donating.

@@

@fig:likely-diffs(B–C) shows the difference in the likelihood of donation for both crackdown conditions across NGO issue areas (H~2a~) and funding sources (H~3a~). Crackdown has almost no effect on the likelihood of donating to human rights NGOs (`r filter(likely, title == "Human rights")$p_neq0`). We hypothesized that the positive crackdown effect would be stronger for humanitarian NGOs, and while crackdown does have a positive effect under the humanitarian frame, the probability this difference is greater than zero is below our 95% threshold (`r filter(likely, title == "Humanitarian assistance")$p_neq0`). We do find some evidence for H~3a~, however. Crackdown has little effect on the likelihood of individual preferences to donate to NGOs receiving government funding (`r filter(likely, title == "Government")$p_neq0`), but it does increase the likelihood of donating to privately-funded NGOs by `r percent_0(filter(likely, title == "Private")$pct_change)` (`r filter(likely, title == "Private")$delta`; `r filter(likely, title == "Private")$p_neq0`).

While the crackdown frame has no effect on the probability of donating to human rights NGOs alone, conditioning this finding on the source of NGO funding reveals competing trends. @fig:likely-diffs(D) shows the difference in donation likelihood across all experimental conditions. When donors know that a human rights organization is primarily government-funded, they are more likely to donate in the *absence* of crackdowns than when the NGO faces legal difficulties (`r filter(likely, title == "Human rights | Government")$delta`; `r filter(likely, title == "Human rights | Government")$p_neq0`). In contrast, when donors know that an organization is privately-funded, they are more likely to donate when there *is* a legal crackdown than when the NGO faces no legal hurdles (`r filter(likely, title == "Human rights | Private")$delta`; `r filter(likely, title == "Human rights | Private")$p_neq0`). For humanitarian assistance NGOs, however, legal crackdowns do not have a significant effect on the likelihood of donation when nesting funding and issue frames.

These results give partial support to H~3a~. The source of funding has little influence on the preferences to donate to humanitarian NGOs, but on average, donors are substantially more likely to want to give to privately-funded human rights NGOs than government-funded organizations. However, the effect of crackdowns reverses across conditions, as donors appear to both punish government-funded human rights NGOs that face legal difficulties and rally behind privately-funded NGOs facing those same hurdles. Respondents described their reasoning for the former as "[i]f the non-profit works in countries where non-profit activities are heavily restricted there's a good amount of waste and shortfalls in their activities and ability to assist others."[^3154467] Respondents also explained their motivation for increased support to privately funded NGOs as "[t]hey [IRC] are doing good work in countries where it is tough for groups like them to operate and they need all the help they can get."[^3025921] Individual donors thus seem to be more willing to support besieged human rights organizations when they are unencumbered by government funds.

@@

Crackdowns do not substantially influence donors' *likelihood* to donate, but they do increase the *amount* of money that respondents are willing to contribute (see @fig:amount-diffs(A)). Informing participants that IRC faces legal hurdles abroad increased donations by `r md_dollar(filter(amount, title == "Crackdown")$mu_diff)` on average, a `r percent_0(filter(amount, title == "Crackdown")$pct_change)` increase from the no crackdown condition (`r filter(amount, title == "Crackdown")$p_neq0`). This provides good evidence for H~1b~—donors appear to give more money to besieged INGOs. In contrast, neither issue area nor funding have an effect on the amount donated (`r filter(amount, title == "Issue")$delta`; `r filter(amount, title == "Issue")$p_neq0`; `r filter(amount, title == "Funding")$delta`; `r filter(amount, title == "Funding")$p_neq0`), indicating that crackdown is a more powerful heuristic for determining private donations.

@@

Combining issue areas and funding sources adds more texture to these results. We previously found that donors seem to (1) punish government-funded human rights NGOs and (2) reward privately-funded human rights NGOs when they face legal issues. This finding also applies to preferences regarding the amount donors donate. Respondents donated `r md_dollar(filter(amount, title == "Human rights | Private")$mu_diff)` more to besieged privately-funded human rights NGOs compared to the no crackdown control condition (`r filter(amount, title == "Human rights | Private")$p_neq0`), increasing their donations by `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)`. Respondents explained that they chose to give additional money to privately-funded human rights NGOs precisely because "the country limits non-profits."[^13685] The punishment mechanism found previously does not translate to the amount donated—while there is a decrease of `r md_dollar(abs(filter(amount, title == "Human rights | Government")$mu_diff))` in average donations to government-funded human rights NGOs facing crackdown, the chance this difference is less than zero is lower than our threshold (`r filter(amount, title == "Human rights | Government")$p_neq0`).
